### 2023-11-01 ncomms/cover.tex

\documentclass[11pt]{article}

\renewcommand\thefigure{L-\arabic{figure}}
\renewcommand\thetable{L-\arabic{table}}
\renewcommand\thepage{L-\arabic{page}}
\renewcommand\theequation{L\arabic{equation}}

\usepackage{color}
\usepackage[normalem]{ulem}% to use \sout in feedback commands
%orange for EI
\definecolor{orange}{rgb}{1,0.5,0}
\newcommand\ei[2]{\sout{#1} \textcolor{orange}{#2}}
\newcommand\eic[1]{\textcolor{orange}{[#1]}}
%pink for KA
\definecolor{pink}{rgb}{1,0.6,0.65}
\newcommand\ka[2]{\sout{#1} \textcolor{pink}{#2}}
\newcommand\kac[1]{\textcolor{pink}{[#1]}}
%purple for AK
\definecolor{purple}{rgb}{0.5,0,1}
\newcommand\ak[2]{\sout{#1} \textcolor{purple}{#2}}
\newcommand\akc[1]{\textcolor{purple}{[#1]}}
%teal for JP
\definecolor{teal}{rgb}{0,0.5,0.8}
\newcommand\jpc[1]{\textcolor{teal}{[#1]}}

\newcommand\vsp{\vspace{2mm}}

\usepackage{fullpage}
\bibliographystyle{apalike}

\usepackage{natbib,amsmath,amsthm}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{amsmath,amssymb}
\usepackage{amsfonts}
\usepackage{natbib}
 \usepackage{url}

 %% code macros
\newcommand\proglang[1]{\textsf{#1}}
\newcommand\code[1]{\texttt{#1}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}
\newcommand\slot[1]{\code{#1}}
\newcommand\class[1]{class `\code{#1}'}
\newcommand\Class[1]{Class `\code{#1}'}

%\newcommand\ifLetter[2]{{#1}}

\parskip 7pt
\parindent 0pt
\newcommand\report[1]{{\color{red} \vspace{1mm}\hspace{0.25in}\parbox{6in}{\em #1}}}
\newcommand\article[1]{{\color{blue} \vspace{1mm}\hspace{0.25in}\parbox{6in}{\em #1}}}
\newcommand\blue[1]{{\color{blue}{#1}}}

%\topmargin -0.3in % for EI dvips

\begin{document}

\rule{0cm}{0.3cm}

%\vspace{-28mm}
\vspace{-8mm}

\includegraphics[width=0.85\textwidth,trim={1.7cm 25.6cm 10cm 1.3cm},clip]{/Users/ionides/Dropbox/computing/latex/umich-stat-letterhead-2019.pdf}

\vspace{5mm}

\begin{flushright}
\today
\end{flushright}

\noindent {\bf Re: ``Machine Learning for Mechanistic Models of Metapopulation Dynamics''}

\vspace{3mm}

Dear Nature Communications editorial team,

\vspace{2mm}

Thank you in advance for your consideration of our manuscript.
We are submitting this as a research article, and indeed the new results reported in the paper are a substantial part of its contribution.
However, the overall goal of this paper is to review and critique previous methodologies while showing how state-of-the-art methods can overcome weaknesses of previous analysis.
The primary role of the new scientific results is to develop a concrete example demonstrating the practical significance of the methodological points.

We are expecting that this topic will be of widespread interest to those scientists whose work involves dynamic interactions between spatially distinct populations. 
Consequently, a high-profile journal such as {\em Nature Communcations} may be an appropriate publication venue.
We look forward to finding out how this work is received at  {\em Nature Communcations}.

Sincerely,

\vspace{-2mm}

\hspace{3cm}
\includegraphics[width=2cm,angle=90]{/Users/ionides/Dropbox/computing/latex/ei-signature1.pdf}

\vspace{-5mm}

Edward Ionides \\
Professor of Statistics\\
University of Michigan\\
ionides@umich.edu

\end{document}