### 2024-02-03 filter_tests/simuplots_filter_tests.Rmd
updated to match arxiv submission

---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r message=FALSE, warning=FALSE, echo=FALSE, fig.width=8, fig.height=10, fig.show='asis'}
library(metapoppkg)
library(ggplot2)
library(ggnewscale)

U <- c(373, 256, 128, 64, 32, 16, 8, 4, 2)

load(file="run_level_2/spatPomp10.rda")
plot_list10 <- list()
obs_counts <- lapply(li23_list, obs)
limit <- log10(max(unlist(obs_counts))+1)
for(i in 1:length(U)) plot_list10[[i]] <- plot_li(x=li23_list[[i]],U=U[i],limit=limit) + theme(legend.position = "none") + ggtitle("filter_tests10")

load(file="run_level_2/spatPomp9.rda")
plot_list9 <- list()
obs_counts <- lapply(li23_list, obs)
limit <- log10(max(unlist(obs_counts))+1)
for(i in 1:length(U)) plot_list9[[i]] <- plot_li(x=li23_list[[i]],U=U[i],limit=limit) + theme(legend.position = "none") + ggtitle("filter_tests9")
for(i in 1:length(U)) print(plot_list10[[i]]/plot_list9[[i]])
```